Species richness and network structure jointly drive total biomass and its temporal stability in fish communities

### 2020-12-07 manuscript/bef_stability/manuscript.Rmd
Introduction reviewed according to elisa comments

Current biodiversity losses and further global change has the potential to affect
ecosystems in complex ways [@loreau_biodiversity_2001; @tilman_biodiversity_2014]. Biodiversity has a
tremendous effect on ecosystem properties because it largely determines both the
amount and the distribution of matter and energy fluxes
within and among ecosystems [@loreau_meta-ecosystems:_2003]. Biodiversity thus affects biomass
production [@tilman_productivity_1996; @cardinale_species_2002;
@grace_integrative_2016], distribution [@thebault_relationship_2006] and
stability [@thebault_trophic_2005; @thebault_relationship_2006;
@tilman_biodiversity_2006]. The relationship between biodiversity and
ecosystem properties is a central topic in ecology [@hector_plant_1999;
@loreau_biodiversity_2001; @hooper_effects_2005; @dunne_network_2006;
@duffy_why_2009; @isbell_forest_quantifying_2018] and our understanding of it is 
crucial to better anticipate the impacts of global change.

@@

Species richness promotes higher biomass production in a variety of ecosystems
and taxons. Indeed species richness generally increases primary productivity and
total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders, [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. Despite the diversity of
ecosystems and taxons studied, the relationship between diversity and ecosystem
functioning remains traditionally focused on single trophic levels, ignoring the
effects of foodweb structure. Theoretical models have recently moved from
single trophic [@tilman_plant_1997;
@loreau_biodiversity_2001] to multitrophic communities [@thebault_food-web_2003;
@poisot_dissimilarity_2012; @schneider_animal_2016; @wang_biodiversity_2018]. 
These models show that both complementarity in the use of resources and selection effects,
the two main mechanisms previously highlighted in single-trophic studies, are
also important to understand diversity effects in foodwebs
[@thebault_food-web_2003; @poisot_dissimilarity_2012; @schneider_animal_2016].
However, they also reveal new mechanisms such as positive effects of 
the number of trophic levels, i.e. associated with higher trophic levels, on ecosystem
functioning [@wang_biodiversity_2018]. Theoretical studies also showed that food
web properties, such as interaction distribution and strength, affect the
distribution of biomass among adjacent trophic levels
[@thebault_relationship_2006; @schneider_animal_2016; @barbier_pyramids_2019].
Despite these theoretical evidence that structure affects biomass, we are
lacking empirical studies that assess jointly how foodweb structure and species
diversity affect total biomass and its production.

---
#### Single and multitrophic communities
---

Further than biomass production, many studies showed that species richness
increases temporal stability of communities, i.e. variation through time, often
quantified by the temporal CV of biomass.  Such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020]. However, this relationship
has been mostly explored in single trophic level systems or without accounting
for foodweb structure, even more so than the relationship between diversity and
biomass production.

Several theoretical studies considering a single trophic level have
clarified which mechanisms generate this positive relationship between species
richness and temporal
stability [@loreau_species_2008; @thibaut_understanding_2013], (1) species
synchrony and (2) species variability. Species synchrony captures to what extent
the fluctuations of abundances or biomasses through time are correlated among
species. A low synchrony in species fluctuations, i.e. asynchrony, is
synonymous of compensatory dynamic, where the decrease in abundance of a species
is compensated by the increase of an other. Asynchrony is generally the main
mechanism which species richness has a stabilizing effect on temporal
stability [@tilman_biodiversity_2006; @olivier_urbanization_2020]. Asynchrony is
expected to increase with the variability of species responses to perturbations. Though most
often positive, the relationship between species diversity and asynchrony has
also been evidenced to be neutral or negative in natural settings
[@tilman_biodiversity_2006; @thibaut_understanding_2013;
@olivier_urbanization_2020; @valencia_synchrony_2020].

In addition to diversity,
foodweb structure is expected to strongly affect the stability of communities
and ecosystems, as outlined by the large body of theoretical studies on this
topic [@may_will_1972; @dunne_network_2006; @neutel_reconciling_2007;
@duffy_why_2009; @stouffer_compartmentalization_2011]. In particular, predation
has cascading effects on both species variability and synchrony
[@teng_dynamics_2004; @shanafelt_stability_2018] and it also affects
synchrony between species.

---
#### Problematic
---

Overall, the evidence suggests that both species richness and network structure
should modulate total biomass and stability of communities. The effects of
species richness and network structure, such as connectance and average trophic
level, are potentially conflicting or synergistic in such a way that they need
to be addressed jointly. Despite the long-standing interest for the links
between foodweb complexity and stability in ecology, we still have very few
evidence of the relationship between network structure and the temporal
stability of multitrophic communities in natural settings.
Although the relationships between diversity, total biomass and its stability
found in experimental and in natural settings can be similar
[e.g., @wu_relationship_2015; @li_relationship_2018], negative or non
significant effects of species richness on total biomass have been found in some
natural communities such as forests [@szwagrzyk_above-ground_2007;
@jacob_productivity_2010; @seidel_relationship_2013] and marine fishes [@maureaud_biodiversityecosystem_2019]. Some authors suggest that
the lack of resource complementarity, environmental harshness or past human
pressures might be responsible for negligible, or negative effects of species
richness [@loreau_biodiversity_2001; @he_global_2013; @wu_relationship_2015;
@maureaud_biodiversityecosystem_2019].

---
#### Environment can mitigate community structure and ecosystem properties 
---

When analyzing data from natural communities, the relationships among species
richness, total biomass and species richness might also arise from effects of
the environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013;
@grace_integrative_2016]. Environment affects organism performance directly, and
thereby, species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020] and temporal
stability [@de_boeck_patterns_2017; @hansen_climate_2013; @bluthgen_land_2016;
@oliver_heterogeneous_2010; @franssen_annual_2011; @friedpetersen_drivers_2020].
For example, larger ecosystem might increase the number of available
resources and thus total biomass [@post_ecosystem_2000; @doi_resource_2009].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the effects of environment and community structure
on total biomass and its stability.

>>>>

Current biodiversity losses and further global change have the potential to
affect ecosystems in complex ways [@loreau_biodiversity_2001;
@tilman_biodiversity_2014]. Biodiversity has indeed a tremendous effect on
ecosystem functions and their stability in response to perturbations
[@hector_plant_1999; @loreau_biodiversity_2001; @hooper_effects_2005;
@dunne_network_2006; @duffy_why_2009], because it largely determines both the
amount and the distribution of matter and energy fluxes within and among
ecosystems [@loreau_meta-ecosystems:_2003]. Understanding the relationship
between biodiversity and ecosystem properties is thus crucial to better
anticipate the impacts of global change.

@@

Biodiversity affects biomass production [@tilman_productivity_1996;
@cardinale_species_2002; @grace_integrative_2016], distribution
[@thebault_relationship_2006] and stability [@thebault_trophic_2005;
@thebault_relationship_2006; @tilman_biodiversity_2006].
Species richness has been shown to promote higher biomass production and  
higher temporal stability of total biomass in a variety of ecosystems
and taxa. 
Indeed species richness generally increases primary productivity and
total biomass in experimental plant communities [@tilman_productivity_1996;
@grace_integrative_2016; @hector_plant_1999; @tilman_biodiversity_2014;
@grace_integrative_2016], but also in briophyte communities
[@mulder_physical_2001], stream feeders, [@cardinale_species_2002] and aquatic
bacterial consumers [@pennekamp_biodiversity_2018]. 
In parallel, such positive relationship between
species richness and temporal stability of biomass has been reported in
experimental grasslands and aquatic bacterial systems
[@tilman_biodiversity_2006; @pennekamp_biodiversity_2018] as well as in natural
settings, such as stream fish communities [@franssen_annual_2011], bird and
butterfly communities [@olivier_urbanization_2020].
Despite the diversity of
ecosystems and taxa studied, the relationship between diversity and ecosystem
functioning in experimental and empirical studies remains mostly focused
on single trophic levels, ignoring the effects of foodweb structure. 
This gap is even more striking for the relationship between diversity and
temporal stability of ecosystem functioning, in spite of long-standing interest
for the links between food web complexity and stability in theoretical ecology
[e.g. @may_will_1972; @dunne_network_2006; @neutel_reconciling_2007;
@duffy_why_2009; @stouffer_compartmentalization_2011].

Overall, the evidence from theoretical models suggests that both species
richness and foodweb structure should modulate ecosystem propertiestotal
biomass and their stability of communities. Complementarity in the use of
resources, a main mechanism previously highlighted in single-trophic studies, is
important to understand the effects of species richness in food webs [@thebault_food-web_2003;
@poisot_trophic_2013; @schneider_animal_2016], but in that case it also depends
on food web connectance [@thebault_food-web_2003; @poisot_trophic_2013].

As classically found in studies considering a single trophic level,
[@tilman_biodiversity_2006; @loreau_species_2008; olivier_urbanization_2020],
asynchrony in species temporal variations is also explaining a significant part
of the stabilizing effects of species richness on total biomass in food webs
[@thebault_trophic_2005]. Indeed, a low synchrony in species fluctuations,
i.e. asynchrony, is synonymous of compensatory dynamics, where the decrease in
abundance of a species is compensated by the increase of an other. However,
higher connectance can in some cases reverse the positive effects of species
richness on temporal stability of total biomass, by leading to a negative
relationship between species richness and species temporal variability
[@thebault_trophic_2005]. In addition, recent models revealed new mechanisms by
which food web structure affects total ecosystem biomass and its temporal
variability. In particular, the number of trophic levels in food webs, which is
associated with the presence of species at higher trophic levels, increases
total primary production [@wang_biodiversity_2018]. The addition of predators has
also cascading effects on both species variability and synchrony at lower
trophic levels [@teng_dynamics_2004; @shanafelt_stability_2018].
Overall, these theoretical studies show that the effects of species richness and
network structure, such as connectance and average trophic level, are
potentially conflicting or synergistic in such a way that they need to be
addressed jointly.

The gap in studies that consider jointly effects of diversity and food web
structure on ecosystem functions can be explained by the challenges associated
with manipulating both species richness and food web complexity in an
experimental context [e.g. @sander_ecological_2017] and by the fact that describing
trophic interactions between species is highly labour intensive. These
challenges are even more acute for the study of the stability of ecosystem
properties, which additionally requires time-series data and longer-term
analyses. In this context, recent developments to predict species trophic
interactions [@gravel_inferring_2013], combined with observational long-term surveys
of ecological communities, can offer an opportunity to investigate the links
between diversity, food web structure and the stability of total community
biomass in natural settings.

Such approach based on data from natural communities can in addition give
insights on the effects of environmental variables on ecosystem properties and
their stability. Environment affects organism performance directly, and thereby,
species abundances, total biomass [@grace_integrative_2016; @woods_testing_2020]
and its temporal stability [@de_boeck_patterns_2017; @hansen_climate_2013;
@bluthgen_land_2016; @oliver_heterogeneous_2010; @franssen_annual_2011;
@friedpetersen_drivers_2020]. For example, larger ecosystem might increase the
number of available resources and thus total biomass [@post_ecosystem_2000;
@doi_resource_2009] while temperature affects ecosystem respiration and biomass
distribution [@yvondurocher_warming_2011; @schwarz_warming_2017]. Environment also
directly affect species richness and food web structure. For instance, larger
ecosystems host food webs with higher diversity and trophic levels
[@mchugh_dual_2010] while temperature can affect food web diversity and
structure in contrasted ways depending on the ecosystem studied
[@ogorman_unexpected_2017; @gauzens_biodiversity_2020]. When analyzing data
from natural communities, the relationships among species richness, total
biomass and its temporal stability might also arise from effects of the
environment on both diversity and ecosystem functions
[@loreau_biodiversity_2001; @he_global_2013; @grace_integrative_2016].
@li_relationship_2018 have notably shown that climate moisture has a positive
direct effect on total biomass, but also indirect through a positive effect on
species richness. Accounting for environment also allow for an overall
comparison of the strength of the effects of environment and community structure
on total biomass and its stability.
